HW3, Ashish EDA

### 2022-09-13 hw3-1/ashish/hw3_test.Rmd
draft

---
title: "HW3, Ashish EDA"
output: html_notebook
---

```{r}
data = read.table('../temps.txt', sep="\t", dec=".", header = TRUE)
data
```

```{r}
data_celsius = 
```

```{r}
CUSUM_ad = function(x,days = data$DAY ,C = 0.30,T = 4){
  mean_x = mean(x)
  sd_x = sd(x)
  C = C * sd_x
  #apply cusum
  x1 = lapply(x, function(xi)  (mean_x - xi - C))
  st_df = data.frame(st=double())
  st = 0
  for(i in 1:length(x)){
    #print (paste(i,class(st),class(x1[[i]])))
    st = max(0, st + x1[[i]])
    st_df[i,] = c(st)
  }
  #Find index where St > T
  st_df$Day = days
  st_df["anamoly"] = 0
  st_df$anamoly[st_df$st > T] = 1
  st_df$xt = x
  #plot x with T and if decrease see
  st_df
}
```

```{r}
check = CUSUM_ad(data$X1996)
#library(ggplot2)
par(mfrow=c(2,1))
plot(check$x)
plot(check$st)
#ggplot(check, aes(x = Day, y = st)) + geom_point() + geom_line() +
#  theme(axis.text.x = element_text(angle = 90))
```

>>>>

---
title: "HW3, Ashish EDA"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(grid)

data = read.table('../temps.txt', sep="\t", dec=".", header = TRUE)
data
```

```{r}
CUSUM_ad = function(x, year, days = data$DAY ,C = 0.30,T = 4, if_c_relative = FALSE){
  mean_x = mean(x)
  sd_x = sd(x)
  if (if_c_relative) {
    C = C * sd_x
  }
  #apply cusum
  x1 = lapply(x, function(xi)  (mean_x - xi - C))
  st_df = data.frame(st=double())
  st = 0
  for(i in 1:length(x)){
    #print (paste(i,class(st),class(x1[[i]])))
    st = max(0, st + x1[[i]])
    st_df[i,] = c(st)
  }
  #Find index where St > T
  st_df$Day = days
  st_df["anamoly"] = 0
  st_df$anamoly[st_df$st > T] = 1
  st_df$xt = x
  #plot x with T and if decrease see
  par(mfrow=c(2,1))
  plot(st_df$x)
  title(paste("Temperature (in Celsius) and S_t (bottom) for year",year), line = -1, outer = TRUE)
  plot(st_df$st)
  pushViewport(viewport())
  grid.lines(x = c(0,1), y = grconvertY(T, "user", "ndc"), gp = gpar(col = "red"))
  change_detected = which.max(st_df$st>T)
  grid.lines(x = grconvertX(change_detected, "user", "ndc"), y = c(0,1), gp = gpar(col = "red"))
  popViewport()
  #return first point where St touches T
  change_length = length(days)-change_detected
  return (c(as.numeric(year),as.numeric(change_detected),as.numeric(change_length)))
}
```

```{r}
CUSUM_ad(data$X1996,"1996", C= 5, T = 50)
#library(ggplot2)
#par(mfrow=c(2,1))
#plot(check$x)
#plot(check$st)
#ggplot(check, aes(x = Day, y = st)) + geom_point() + geom_line() +
#  theme(axis.text.x = element_text(angle = 90))
```

```{r}
### Transform to celsius
temp_cols = names(data)[!(names(data) %in% c("DAY"))]
#Transform to celsius
data_celsius = data
data_celsius[temp_cols] <- lapply(data_celsius[temp_cols], function(f) (f-32)*5/9)
```

```{r}
CUSUM_ad(data_celsius[,"X1996"],1996, C = 3, T = 25)
```

```{r}
### Run Cusum over data_celsius for all the years
mean(unlist(lapply(temp_cols, function(col) 
  sd(data_celsius[,col])
  )))
```

```{r}
### Run Cusum over data_celsius for all the years
winter_start_days = lapply(temp_cols, function(col) 
  CUSUM_ad(x=data_celsius[,col], year = substr(col,2,5), C = 6, T = 25)
  )

# barplot(sapply(winter_start_days,"[[",1),
# main = "Variation of Winter Start from 1996 to 2015",
# xlab = "Year",
# ylab = "Days since Jul 1, when winter change detected",
# names.arg = temp_cols,
# col = "darkred",
# horiz = FALSE)
```

```{r}
### Run Cusum over data_celsius for all the years
winter_start_days = lapply(temp_cols, function(col) 
  CUSUM_ad(x=data_celsius[,col], year = substr(col,2,5), C = 4, T = 25)
  )
```

```{r}
change_df <- as.data.frame(do.call(rbind, winter_start_days))
colnames(change_df) = c("year","winter_start","winter_length")
change_df["year_bucket"] = "a.1996_2000"
change_df[change_df$year>=2000,"year_bucket"] = "b.2000_2004"
change_df[change_df$year>=2004,"year_bucket"] = "c.2004_2008"
change_df[change_df$year>=2008,"year_bucket"] = "d.2008_2012"
change_df[change_df$year>=2012,"year_bucket"] = "e.2012_2016"
change_df
```

```{r}
aggregate(cbind(winter_start,winter_length) ~ year_bucket, change_df, mean)
```

```{r}
barplot(sapply(winter_start_days,"[[",2),
main = "Variation of Winter Start from 1996 to 2015",
xlab = "Year",
ylab = "Days since Jul 1, when winter change detected",
names.arg = temp_cols,
col = "darkred",
horiz = FALSE)
```

```{r}
plot(sapply(winter_start_days,"[[",1))
axis(1, at=1:length(temp_cols),labels=temp_cols, las=3)
```
